Compiling avoidable deaths codes (as coded by NRS), 2020

### 2022-05-24 compile_avoidable_death_codes_2020.Rmd
Fixed error in parsing age variable where spaces and some kind of invisible trailing character were erroneously included

```{r}
avoidable_deaths <-
  avoidable_deaths_raw %>%
  drop_na(condition_group_and_cause) %>%  # remove blank rows used for blank space in the excel file
  mutate(
    group = if_else(condition = is.na(icd_10_codes), true = condition_group_and_cause, false = NA_character_),
    original_codes = icd_10_codes
    ) %>%
  fill(group, .direction = "down") %>%  # forward-fill group
  drop_na(icd_10_codes) %>%
  separate_rows(icd_10_codes, sep=",\\s*", ) %>%  # separate comma-separated values into rows so we have one entry per row
  extract(col = icd_10_codes, into = "exclusions", regex = "\\(excl\\.\\s*(.+)\\)", remove = FALSE) %>%  # extract the exclusions and put them in a separate column
  mutate(
    icd_10_codes = str_remove_all(string = icd_10_codes, pattern = "\\s*\\(excl.+\\)"),  # remove the exclusions from the code ranges
    icd_10_codes = map_chr(.x = icd_10_codes, .f = ~parse_icd_10_range(.x))
  ) %>%
  separate_rows(icd_10_codes, sep=",\\s*") %>%  # separate comma-separated values again now that we've converted them from the ranges
  extract(col = age, into = c("age_min", "age_max"), regex = "(\\d+)[^0-9]+(\\d+)", remove = TRUE) %>%
  mutate(
    treatable = case_when(  # make more explicit the treatable/preventable distinction
      str_detect(treatable, pattern = "50") ~ 0.5,
      str_detect(treatable, pattern = "\\•") ~ 1,
      TRUE ~ 0
    ),
    preventable = case_when(
      str_detect(preventable, pattern = "50") ~ 0.5,
      str_detect(preventable, pattern = "\\•") ~ 1,
      TRUE ~ 0
    ),
  ) %>%
  relocate(group, condition_group_and_cause, original_codes)

@@

>>>>

```{r}
avoidable_deaths <-
  avoidable_deaths_raw %>%
  drop_na(condition_group_and_cause) %>%  # remove blank rows used for blank space in the excel file
  mutate(
    group = if_else(condition = is.na(icd_10_codes), true = condition_group_and_cause, false = NA_character_),
    original_codes = icd_10_codes
    ) %>%
  fill(group, .direction = "down") %>%  # forward-fill group
  drop_na(icd_10_codes) %>%
  separate_rows(icd_10_codes, sep=",\\s*", ) %>%  # separate comma-separated values into rows so we have one entry per row
  extract(col = icd_10_codes, into = "exclusions", regex = "\\(excl\\.\\s*(.+)\\)", remove = FALSE) %>%  # extract the exclusions and put them in a separate column
  mutate(
    icd_10_codes = str_remove_all(string = icd_10_codes, pattern = "\\s*\\(excl.+\\)"),  # remove the exclusions from the code ranges
    icd_10_codes = map_chr(.x = icd_10_codes, .f = ~parse_icd_10_range(.x))
  ) %>%
  separate_rows(icd_10_codes, sep=",\\s*") %>%  # separate comma-separated values again now that we've converted them from the ranges
  extract(col = age, into = c("age_min", "age_max"), regex = "(\\d+)[^0-9]+(\\d+)", remove = TRUE, convert = TRUE) %>%
  mutate(
    treatable = case_when(  # make more explicit the treatable/preventable distinction
      str_detect(treatable, pattern = "50") ~ 0.5,
      str_detect(treatable, pattern = "\\•") ~ 1,
      TRUE ~ 0
    ),
    preventable = case_when(
      str_detect(preventable, pattern = "50") ~ 0.5,
      str_detect(preventable, pattern = "\\•") ~ 1,
      TRUE ~ 0
    ),
  ) %>%
  relocate(group, condition_group_and_cause, original_codes)

@@

## Checking parsing of age

There was previously an error with this, with some ages having a leading space, and a trailing invisible character. This is slightly overkill to check since the minimum and maximum ages are the same for all conditions, but it may change in the future! I fixed this by using `extract()` to parse the digits from the entries rather than just splitting them on the hyphen symbol.

```{r}
avoidable_deaths_raw %>% 
  count(age) %>%
  kable(caption = "The different possible entries for 'age' in the raw spreadsheet. Note how there are two identical-looking entries, probably due to an invisible character.")

avoidable_deaths%>% 
  count(age_min, age_max) %>%
  kable(caption = "The different possible entries for 'age' in the clean data after parsing.")
```
